---
output: github_document
bibliography: README.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- [![Travis-CI Build Status](https://travis-ci.org/sflippl/vdem.tectr.svg?branch=master)](https://travis-ci.org/sflippl/vdem.tectr)[![Coverage Status](https://img.shields.io/codecov/c/github/sflippl/vdem.tectr/master.svg)](https://codecov.io/github/sflippl/vdem.tectr?branch=master) -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# vdem.tectr

vdem.tectr provides access to Version 8 of the [V-Dem dataset](v-dem.net) [@vdem2018, @Pemstein2018]. It contains:

* `df_vdem`, the imported V-Dem dataset
* `vdem_spatial` and `vdem_geocode`, which geocode the countries in the V-Dem dataset
* `vdem`, `df_vdem` equipped with a metaframe which uses effective explicitness from the package [`tectr`](github.com/sflippl/tectr) to asssist in visualizations and descriptions of variables.

## Installation

You can install vdem.tectr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("sflippl/vdem.tectr")
```
